---
output: 
  html_document:
    css: assessment_platform.css

params:
  compartment: must_supply
  series: must_supply
---

```{r pagetitle, include = FALSE}
page_title <- params$series
page_title <- gsub("_", " ", page_title)
page_title <- stringi::stri_trans_general(page_title, "LATIN-ASCII")
```

---
pagetitle: `r page_title`
---


```{r load_packages, include = FALSE}
require(lattice)
require(grid)
require(knitr)

# will also require libraries stringi and stringr to be installed, but not loaded

knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
options(width = 110)
```


```{r setup_data, echo = FALSE}

#### get key structures ####

assessment_object <- get(paste0(params$compartment, "_web"))

determinands <- assessment_object[["determinands"]]

assessment_object <- assessment_object[["assessment"]]


##### get data and assessments in the station / determinand group ####

# biota: need further subdivision by species and matrix (contaminants only)
# convert detGroup to character just to be on safe side!

assessment_object$timeSeries <- within(assessment_object$timeSeries, {
  detGroup <- as.character(detGroup)
})

var_id <- c("station", "detGroup")
if (params$compartment == "biota") {
  var_id <- c(var_id, "species")
  group_id <- assessment_object$timeSeries[params$series, "detGroup"]
  if (!group_id  %in% c("Effects", "Imposex", "Metabolites"))
    var_id <- c(var_id, "matrix")
  species_id <- assessment_object$timeSeries[params$series, "species"]
  if (get.info("species", species_id, "family") %in% "Mammal")
    var_id <- c(var_id, "AMAP_group")
}

assessment_object <- within(assessment_object, {
  timeSeries$multi_series <- do.call("paste", timeSeries[var_id])
})

id <- assessment_object$timeSeries[params$series, "multi_series"]
assessment_object <- ctsm.subset.assessment(assessment_object, multi_series %in% id)

# sort, using determinands, to ensure everything is in the 'correct' presentational order

assessment_object <- within(assessment_object, {
  timeSeries <- timeSeries[order(match(timeSeries$determinand, determinands)), ]
})


#### get data, assessment and info for params$series ####

ok <- with(assessment_object$data, seriesID %in% params$series)
data <- assessment_object$data[ok, ]

assessment <- assessment_object$assessment[[params$series]]

info <- with(assessment_object, c(info, timeSeries[params$series, ]))

# determinand info

info$group <- unique(as.character(data$group))
info$det_name <- get.info("determinand", info$determinand, "common.name")
if (info$group %in% c("Metals", "PAH_parent", "Organotins"))
  info$det_name <- tolower(info$det_name)

# species info

if (info$compartment %in% "biota") {
  info$species_name <- get.info("species", info$species, "common.name")
  ok <- sapply(
    c("European", "Manila", "Atlantic", "Mediteranean", "Baltic", "Eurasian"), 
    grepl, 
    info$species_name
  )
  if (!any(ok)) info$species_name <- tolower(info$species_name)
  
  info$species_group <- get.info("species", info$species, "family")
  info$species_group <- as.character(info$species_group)
}

if (info$compartment %in% "biota")
  info$AMAP_group <- gsub("_", " ", info$AMAP_group)


if (info$group %in% c("Effects", "Imposex")) {
  txt_concentration <- "level"
} else {
  txt_concentration <- "concentration"
}

txt_concentrations <- paste0(txt_concentration, "s")

txt_measurements <- switch(
  txt_concentration, 
  concentration = "concentration measurements",
  "measurements"
)
  
txt_compounds <- if (info$group %in% c("Imposex", "Effects", "Metabolites")) {
  "responses"
} else {
  "compounds"
}
  

info$distribution <- get.info("determinand", info$determinand, "distribution")


# station info

if (is.na(info$stationName)) info$stationName <- info$station

# matrix info

# can be multiple matrices for sediment, in which case take most common
# should only be a single matrix for other compartments

info$matrix <- unique(as.character(data$matrix))

if (params$compartment != "sediment" & length(info$matrix) > 1) 
  cat("Error: multiple matrices - need to investigate")

if (length(info$matrix > 1)) {
  wk <- table(as.character(data$matrix))
  info$matrix <- names(wk[which.max(wk)])
}
 
info$matrix_name <- get.info("matrix", info$matrix, "name")

info$matrix_name <- switch(
  info$matrix_name, 
  "erythrocytes (red blood cells in vertebrates)" = "red blood cells",
  "egg homogenate of yolk and albumin" = "egg yolk and albumin",
  "hair/fur" = "hair",
  info$matrix_name
)


# sex info

if (info$compartment == "biota")
  info$sex_name <- switch(as.character(info$sex), M = "male", F = "female")


# units 

info$unit <- as.character(get.info("determinand", info$determinand, "unit", info$compartment))

info$basis <- as.character(info$basis)

args.list <- list(units = info$unit, basis = info$basis, html = TRUE, compartment = info$compartment)

is_normalised <- info$compartment == "sediment" & !info$region %in% c("Iberian Sea", "Gulf of Cadiz")
if (is_normalised) {
  extra.text <- paste("normalised to", switch(info$group, Metals = "5% aluminium", "2.5% organic carbon"))
  args.list <- c(args.list, extra.text)
}  

info$unit_text <- do.call(label.units, args.list)



is_AC <- length(na.omit(assessment$AC))

if (is_AC)
  AC_id <- names(na.omit(assessment$AC))



# is there a non-parametric test of status

is_nonparametric <- 
  if (info$group %in% "Imposex") {
    FALSE
  } else {
    is_AC && !is.na(assessment$summary[[paste0(AC_id[1], "below")]])
  }


# is there a parametric model fit

is_pred <- "pred" %in% names(assessment)



#### get data, assessment and info for multiplots ####

info_multi <- info

info_multi <- within(info_multi, {
  seriesID <- row.names(assessment_object$timeSeries)
  determinand <- as.character(assessment_object$timeSeries$determinand)
  names(determinand) <- seriesID
})

# get series names for labelling plots - usually just determinand, but could be more complex if e.g.
# measured in multiple tissues, or when dealing with biological effects

info_multi$plotNames <- with(info_multi, list(data = determinand, assessment = determinand))

if (any(duplicated(info_multi$plotNames))) {
  dups <- with(info_multi, duplicated(determinand) | duplicated(determinand, fromLast = TRUE))
  info_multi$plotNames$data[dups] <- paste(
    info_multi$determinand, assessment_object$timeSeries$level6name, sep = "\n")[dups]
  info_multi$plotNames$assessment[dups] <- paste(
    info_multi$determinand, assessment_object$timeSeries$level6name)[dups]
  if (any(duplicated(info_multi$plotNames$data))) 
    cat("Error: duplicate plotting names - need to extend coding")
}
  
info_multi$matrix <- with(assessment_object$data, tapply(matrix, seriesID, function(x) {
  out <- table(as.character(x))
  names(out[which.max(out)])
}))

info_multi <- within(info_multi, {
  group <- get.info("determinand", determinand, "group", compartment)
  group <- as.character(group)
})

```


```{r intro_txt, include = FALSE}
intro_txt <- switch(
  info$compartment, 
  biota = paste(info$species_name, info$matrix_name), 
  info$compartment
)

if (info$compartment %in% "biota" && info$species_group %in% "Mammal") 
  intro_txt <- paste(tolower(info$AMAP_group), intro_txt)

if (info$compartment %in% "biota" && info$determinand %in% "EROD")
  intro_txt <- paste(info$sex_name, intro_txt)

intro_AMAP <- info$compartment %in% "biota" && 
  ((info$species_group %in% "Mammal" & info$station %in% "Svalbard") | 
     (info$species_group %in% "Bird" & info$station %in% "Svalbard-Kongsfjorden area"))

if (info$compartment %in% "biota" && info$species_group %in% "Mammal" && info$station %in% "Svalbard")
  intro_txt2 <- "The data were kindly provided by the Norwegian Polar Institute (in accordance with CC-BY 4.0 licence). "

if (info$compartment %in% "biota" && info$species_group %in% "Bird" && info$station %in% "Svalbard-Kongsfjorden area")
  intro_txt2 <- paste(
    "These data were kindly provided by Olivier Chastel (CEBC), Børge Moe (NINA), ",
    "Geir Wing Gabrielsen (NPI), Jan Ove Bustnes (NINA) and Claus Bech (NTNU). ", 
    "The work was supported by the French Polar Institute and the The Norwegian Research Council."
  )


AC_trial <- (info$group %in% "PBDEs" && is_AC) | 
  (params$compartment %in% "sediment" && info$group %in% "Organotins" && is_AC)
intro_txt3 <- "The assessment criteria are being used on a trial basis. "
```

### Assessment plots and statistical analysis {.tabset}

This report provides details of the assessment of `r paste(info$det_name, txt_concentrations)` in `r intro_txt` at station `r info$stationName`. `r if (intro_AMAP) intro_txt2``r if (is_normalised) paste0("Concentrations are ", extra.text, " unless otherwise stated. ")``r if (AC_trial) intro_txt3`

<ul>
<li class = "gap">Timeseries metadata.  Key information about the timeseries.</li>

```{r txt_variables, include = FALSE}
txt_trend1 <- paste(
  "The",
  if (is_pred) "trend is" else "points are", 
  "plotted on the log scale, with the axis labels back-transformed for ease of interpretation."
)

txt_trend2 <- switch(
  info$distribution,
  lognormal = paste0(
    "The points are the annual medians of the individual log ", txt_measurements, 
    ", with solid circles denoting uncensored values and < denoting less-than values."
  ),
  "The points are the annual means of the individual measurements."
)
  
txt_trend3 <- switch(
  info$distribution, 
  lognormal = paste0(
    "The points are the individual ", txt_measurements, 
    ", with + denoting uncensored values, ", 
    "d or q denoting values below the limit of detection or quantification and ",
    "< denoting other less-than values."
  ), 
  "The points are the individual measurements."
)

txt_AC <- paste(
  "The assessment", 
  if (is_AC && length(AC_id) == 1) "criterion is" else "criteria are", 
  "indicated on the right hand side of the plot and by horizontal dashed lines (if within the range of the data)."
)
```

```{asis, eval = is_pred}
<li class = "gap">Assessment plot.  The fitted trend (solid line) with pointwise 90% confidence bands (grey shaded areas).  `r if (info$distribution == "lognormal") txt_trend1` `r txt_trend2` `r if (is_AC) txt_AC`</li>
<li class = "gap">Trend with data.  The fitted trend with pointwise 90% confidence bands.  `r txt_trend3`</li>
```  

```{asis, eval = !is_pred}
<li class = "gap">Assessment plot. `r txt_trend2` `r if (info$distribution == "lognormal")  txt_trend1` However, there are too few years, or too many less-than values, to fit a model to these data.  `r if (is_AC) txt_AC`</li>
<li class = "gap">Trend with data.  `r txt_trend3`</li>
```  


```{asis, eval = info$compartment %in% "biota"}
<li class = "gap">Auxiliary data.  The individual `r txt_measurements` with supporting information: mean length of the individuals in the sample; dry weight (%) of the sample; lipid weight (%) of the sample.</li>
```

```{asis, eval = info$compartment %in% "sediment"}
<li class = "gap">Auxiliary data.  The individual concentration measurements`r if (is_normalised) paste0(", both non-normalised and ", extra.text, ",")` with supporting information: aluminium content (%) of the sample; `r if (info$group %in% "Metals") "lithium" else "carbon"` content (%) of the sample.</li>
```

```{asis, eval = info$compartment %in% "water"}
<li class = "gap">Auxiliary data.  No auxiliary data are currently plotted.</li>
```


```{asis, eval = info$purpose %in% "AMAP"}
<li class = "gap">Stable isotope data.  The individual `r txt_measurements` with supporting information: the mean length of the individuals in the sample; the isotope ratio 13C:12C and the isotope ratio 15N:14N.</li>
```



```{r txt_related_compounds, include = FALSE}
txt_related1 <- switch(
  info$group, 
  Imposex = "There are no related responses.",
  Effects = paste(
    "Assessment plots for all the biological effects (other than imposex and PAH metabolites) measured in", 
    info$species_name, "at this station."
  ),
  Metabolites = paste(
    "Assessment plots for all the PAH metabolites measured in", 
    info$species_name, "at this station."
  ),
  paste0(
    "Assessment plots for all the compounds in the same chemical group measured ", 
    if (info$compartment == "biota") paste0("in ", info$species_name, " ", info$matrix_name, " "),
    "at this station."
  )
)

txt_related2 <- switch(
  info$group, 
  Imposex = "There are no related responses.",
  Effects = paste(
    "A scatter plot matrix of all the individual biological effects measurements", 
    "(other than imposex and PAH metabolites) in", 
    info$species_name, "at this station."
  ),
  Metabolites = paste(
    "A scatter plot matrix of all the individual PAH metabolite measurements in", 
    info$species_name, "at this station."
  ),
  paste0(
    "A scatter plot matrix of all the individual concentration measurements for all the ",
    "compounds in the same chemical group ",
    if (info$compartment == "biota") paste0("in ", info$species_name, " ", info$matrix_name, " "),
    "at this station."
  )
)
```

<li class = "gap">Assessment (related `r txt_compounds`).  `r txt_related1`</li>
<li class = "gap">Data (related `r txt_compounds`). `r txt_related2`</li>

```{asis, eval = is_pred}
<li class = "gap">Statistical analysis. This tab summarises the fits of models with different amount of smoothing. The final choice is the model with the minimum AICc (with the caveat that a linear model is always preferred to a mean model if there are sufficient years of data). The tab also gives the contrast between the fitted values at the start and end of the timeseries, and over the most recent twenty years.  `r if(is_AC) "Finally, the tab assesses status by comparing the fitted value at the end of the timeseries to the assessment criteria."`</li>
```

```{asis, eval = !is_pred & !is_nonparametric}
<li class = "gap">Statistical analysis. No statistical analysis was done on these data.</li>
```

```{asis, eval = !is_pred & is_nonparametric}
<li class = "gap">Statistical analysis. This tab gives the results of a non-parametric status assessment.</li>
```


</ul>


<br>

#### Timeseries metadata

<br>

<ul>
<li>OSPAR region: `r info$OSPARregion`</li>
<li>Region: `r info$region`</li>
<li>Station code: `r info$station`</li>
<li>Station name: `r info$stationName`</li>
<li>Station latitude: `r format(info$latitude, digits = 2, nsmall = 2)`</li>
<li class = "gap">Station longitude: `r format(info$longitude, digits = 2, nsmall = 2)`</li>
<li>Compartment: `r info$compartment`</li>

```{asis, eval = info$compartment %in% "biota"}
<li>Species latin name: `r info$species`</li>
<li>Species common name: `r info$species_name`</li>  
```

```{asis, eval = info$compartment %in% "biota" && info$species_group %in% "Mammal"}
<li>Sex / age group: `r tolower(info$AMAP_group)`</li>
```

```{asis, eval = info$determinand %in% "EROD"}
<li>Sex: `r info$sex_name`</li>
```

<li class = "gap">`r switch(info$compartment, biota = "Tissue", "Matrix")`: `r info$matrix_name`</li>
<li>Determinand code: `r info$determinand`</li>
<li>Determinand name: `r info$det_name`</li>

```{asis, eval = info$detGroup %in% "PAH metabolites"}
<li>Method of chemical analysis: `r info$metoa`</li>
```

<li class = "gap">Units: `r info$unit_text`</li>
<li>Extraction: `r info$extraction`</li>
</ul>

<br>




#### Assessment plot

```{r assessment_plot, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.data(data, assessment, info, type = "assessment", xykey.cex = 1.4)
```


#### Trend with data

```{r data_plot, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.data(data, assessment, info, type = "data", xykey.cex = 1.4)
```


#### Auxiliary data

```{r, include = FALSE}
ok <- params$compartment %in% c("biota", "sediment")
```

```{asis, eval = !ok}
<br> 
No auxiliary variables currently plotted.
```

```{r auxiliary_data, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
auxiliary_id <- "default"
if (info$compartment %in% "sediment" & info$group %in% "Metals") 
  auxiliary_id <- c("value", "concentration", "AL", "LI")

plot.auxiliary(data, info, auxiliary_id = auxiliary_id, xykey.cex = 1.2)
```



<!-- #### Stable isotope data -->

<!-- ```{r isotope_data, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7} -->
<!-- wk <- switch(info$species_group, Mammal = "LIPIDWT%", "LNMEA") -->
<!-- plot.auxiliary(data, info, auxiliary_id = c("concentration", wk, "C13D", "N15D"), xykey.cex = 1.2) -->
<!-- ``` -->



#### Assessments (related `r txt_compounds`)

```{r, include = FALSE}
ok <- ! info$detGroup %in% "Imposex"
```

```{asis, eval = !ok}
<br>
No related responses assessed.
```

```{r multi_assessment, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.multiassessment(
  assessment_object$data, assessment_object$assessment, info_multi, type = "assessment")
```


#### Data (related `r txt_compounds`)

```{r, include = FALSE}
ok <- ! info$detGroup %in% "Imposex"
```

```{asis, eval = !ok}
<br>
No related responses assessed.
```

```{r multi_data, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.multidata(assessment_object$data, info_multi)
```


#### Statistical analysis

<br>

**Trend assessment**

```{r trend_ok, include = FALSE}
anova <- assessment$anova
coefficients <- assessment$coefficients
if (info$detGroup == "Imposex") {
  anova_ok <- !is.null(coefficients) | !is.null(anova)
} else {
  anova_ok <- !is.null(anova) && nrow(anova) > 1
}
```


```{asis, eval = !anova_ok & !info$group %in% c("Effects", "Imposex")}
There are too few years, or too many less-than values, for a trend assessment.
<br>
```

```{asis, eval = !anova_ok & info$group %in% "Effects"}
There are too few years for a trend assessment.
<br>
```

```{asis, eval = !anova_ok & info$group %in% "Imposex"}
There is no trend assessment because there are too few years (or occasionally because all the measurements are zero).
<br>
```


```{r, eval = anova_ok & info$detGroup != "Imposex", child = "assessment_platform_trend.Rmd"}
```

```{r, eval = anova_ok & info$detGroup == "Imposex", child = "assessment_platform_trend_imposex.Rmd"}
```

<br>

**Status assessment**

```{r status_ok, include = FALSE}

if (info$detGroup == "Imposex") {
  nyear <- assessment$summary$clLY
  nyear_ok <- !is.na(nyear)
} else {
  nyear <- assessment$summary$nyfit
  nyear_ok <- !is.null(nyear) && nyear >= 3
}

AC_ok <- !all(is.na(assessment$AC))
status_param <- nyear_ok && AC_ok

status_ok <- status_param | is_nonparametric

if (is_nonparametric) {
  is_below <- assessment$summary[paste0(AC_id, "below")]
  is_below <- unlist(is_below)
  is_below <- is_below %in% "below"
  names(is_below) <- AC_id
  
  EAC_id <- setdiff(AC_id, "BAC")
  
  if ("BAC" %in% AC_id & is_below["BAC"]) 
    status_txt <- "significantly below the BAC (p < 0.05)."
  else if ("BAC" %in% AC_id & length(AC_id) == 1)
    status_txt <- "not significantly below the BAC (p > 0.05)."
  else if (any(is_below) & length(AC_id) == 2)
    status_txt <- paste(
      "significantly below the", EAC_id, 
      "(p < 0.05), but not significiantly below the BAC (p > 0.05)."
    )
  else if (any(is_below) & length(AC_id) == 1)
    status_txt <- paste("significantly below the", EAC_id, "(p < 0.05).")
  else 
    status_txt <- paste("not significantly below the", EAC_id, "(p > 0.05).")
}

```


```{asis, eval = !status_ok}
`r if (!AC_ok) "There are no assessment criteria so no status assessment is possible." else if (!nyear_ok) "There are insufficient data to make a status assessment."`
<br>
```

```{asis, eval = !status_param & is_nonparametric}
A sign test based on the last five monitoring years shows that the mean `r txt_concentration` was `r status_txt`
<br>
```

```{r, eval = status_param & info$detGroup != "Imposex", child = "assessment_platform_status.Rmd"}
```

```{r, eval = status_param & info$detGroup == "Imposex", child = "assessment_platform_status_imposex.Rmd"}
```


<br>

```{r interpretation_setup, include = FALSE}
format_p <- function(p) {
  if (p >= 0.0001) {
    p <- round(p, 4)
    paste0("(p = ", format(p, scientific = FALSE, nsmall = 4), ")")
  } else {
    "(p < 0.0001)"
  }
}
```


```{r interpretation_trend, eval = anova_ok & !info$group %in% c("Effects", "Imposex"), include = FALSE}
wk <- assessment$summary
wk_year <- unique(assessment$data$year)
wk_n <- length(wk_year)
wk_seq <- length(wk_year) == max(wk_year) - min(wk_year) + 1
wk_recent <- info$maxYear - info$recent.trend + 1


change_txt <- function(type = c("whole", "recent")) {
  p <- switch(type, whole = wk$pltrend, wk$prtrend)
  trend <- switch(type, whole = wk$ltrend, wk$rtrend)
  paste0(
    stringr::str_to_sentence(txt_concentrations),
    " at the end of the time series were ", 
    if (p < 0.05) paste("signficantly", if (trend > 0) "higher" else "lower", "than those ") else "not significantly different from those ",
    switch(type, whole = "at the start of the time series", paste("in", wk_recent)), 
    " ", format_p(p), ". "
  )
}

trend_description <- 
  if (wk$p_overall >= 0.05) {
    paste0("There is no significant temporal trend in the time series ", format_p(wk$p_overall), ".")
  } else if (is.na(wk$p_nonlinear)) {
    paste0(
      "There is a significant log-linear trend in the time series ", 
      format_p(wk$p_linear), 
      ". ",
      stringr::str_to_sentence(txt_concentrations),
      " have ", 
      if (wk$ltrend > 0) "increased" else "decreased", 
      " by an estimated ", 
      round(100 * (exp(wk$ltrend / 100) - 1), 1), 
      "% per year over the course of the time series."
    )
  } else {
    paste0(
      "There is a significant temporal trend in the time series ", 
      format_p(wk$p_overall), 
      ". The trend is nonlinear ", 
      format_p(wk$p_nonlinear), " 
      so the shape of the trend must be assessed visually. ", 
      change_txt("whole"), 
      if (assessment$contrasts["whole", "start"] < wk_recent) change_txt("recent")
    )
  }
```


```{r interpretation_status, eval = status_ok & !info$group %in% c("Effects", "Imposex"), include = FALSE}

status_txt_health <- if ("MPC" %in% AC_id) "human health." else "marine life."

if (is_nonparametric) {
  status_txt2 <- 
    if ("BAC" %in% AC_id & is_below["BAC"]) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore at background levels."
      )
    } else if ("BAC" %in% AC_id & length(AC_id) == 1) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore above background levels, but, as there is not an environmental threshold,", 
        "it is not possible to say whether the",
        txt_concentrations, 
        "would harm", 
        status_txt_health
      )
    } else if (any(is_below) & length(AC_id) == 2) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore above background levels but unlikely to harm", 
        status_txt_health
      ) 
    } else if (any(is_below) & length(AC_id) == 1) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore unlikely to harm", 
        status_txt_health
      ) 
    } else {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "could therefore harm",
        status_txt_health
      )
    }
}

# have to update status_txt if done a parametric test - this should be rationalised in a single status file

if (status_param) {
  EAC_id <- setdiff(AC_id, "BAC")
  
  AC_p <- rv[[p_txt]]
  names(AC_p) <- row.names(rv)

  status_txt <-
    if ("BAC" %in% AC_id & AC_p["BAC"] < 0.05) {
      paste0("significantly below the BAC ", format_p(AC_p["BAC"]), ".")
    } else if ("BAC" %in% AC_id & length(AC_id) == 1) {
      paste0("not significantly below the BAC ", format_p(AC_p["BAC"]), ".")
    } else if (any(AC_p < 0.05) & length(AC_id) == 2) {
      paste0(
        "significantly below the ", EAC_id, " ", format_p(AC_p[EAC_id]),
        ", but not significiantly below the BAC ", format_p(AC_p["BAC"]), "."
      )
    } else if (any(AC_p < 0.05) & length(AC_id) == 1) {
      paste0("significantly below the ", EAC_id, " ", format_p(AC_p[EAC_id]), ".")
    }    
  else {
      paste0("not significantly below the ", EAC_id, " ", format_p(AC_p[EAC_id]), ".")
    }
  
  status_txt2 <- 
    if ("BAC" %in% AC_id & AC_p["BAC"] < 0.05) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore at background levels."
      )
    } else if ("BAC" %in% AC_id & length(AC_id) == 1) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore above background levels, but, as there is not an environmental threshold,", 
        "it is not possible to say whether the",
        txt_concentrations, 
        "would harm", 
        status_txt_health
      )
    } else if (any(AC_p < 0.05) & length(AC_id) == 2) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore above background levels but unlikely to harm", 
        status_txt_health
      ) 
    } else if (any(AC_p < 0.05) & length(AC_id) == 1) {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "were therefore unlikely to harm", 
        status_txt_health
      ) 
    } else {
      paste(
        stringr::str_to_sentence(txt_concentrations),
        "could therefore harm",
        status_txt_health
      )
    }

}

status_description <- paste(
  stringr::str_to_sentence(txt_concentrations),
  "at the end of the time series were",
  status_txt, 
  status_txt2
)
 

```




```{asis, eval = (anova_ok | status_ok) & !(info$group %in% c("Effects", "Imposex"))}
**Interpretation**

`r if (anova_ok) trend_description`

`r if (status_ok) status_description`

<br>

```




```{asis, include = FALSE, eval = FALSE}
The lowest detectable annual increase in the time series (two-sided test, power = 80%, size = 5%) is `r wk$dtrend_1`%. `r if (!wk_seq) paste0("Had the data been collected sequentially (rather than with gaps between years) the lowest detectable annual increase would be ", wk$dtrend_2, "%.")`  The lowest detectable annual increase with 10 years of `r if (!wk_seq) "sequential"` monitoring is `r wk$dtrend_3`%.

Given the variability in the data, `r wk$dyear` years of `r if (!wk_seq) "sequential"` monitoring would be required to detect an annual increase of 10% with 80% power (two-sided test, size = 5%).

The power of the time series to detect an annual increase of 10% is `r wk$dpower_1`% (two-sided test, size = 5%).  `r if (!wk_seq) paste0("With sequential monitoring, the power would be ", wk$dpower_2, "%")`.  With 10 years of `r if (!wk_seq) "sequential"` monitoring, the power would be `r wk$dpower_3`%.
```




```{r tidy_up, include = FALSE}
rm(list = intersect(
  objects(),
  c("AC_ok", "anova", "anova_ok", "assessment", "assessment_object", "change_txt", "coefficients",
    "contrasts", "data", "determinands", "diff_txt", "dups", "good_status", "id", "info", "info_multi",
    "is_lognormal", "nyear", "nyear_ok", "ok", "p_txt", "pred", "rv", "status_ok", "wk")
))
```
